\papertitle

### 2025-04-17 paper/main.tex

    \title{\papertitle}
    \author[]{Redacted for review}%{\myname}
    \affil[]{Redacted for review}%{\myaffiliation}
    \date{ }%{\paperdate}

@@

        Travel time prediction is central to accessibility analyses, sustainable transportation infrastructure provision, and active transportation interventions. However, calculating accurate travel times, especially for driving, requires either extensive technical capacities and bespoke data, or sources like the Google Maps API that quickly becomes prohibitively expensive to analyze thousands or millions of trips necessary for metropolitan-scale planning. These obstacles make it difficult for researchers, practitioners, and community advocates to develop an evidence base for more sustainable transportation and equitable accessibility. This paper develops a free, open-source minimally-congested driving time prediction model with minimal cost, data, and computational requirements. It trains and tests this model using the Los Angeles, California urban area as a case study by calculating naïve travel times then developing a predictive random forest regression model of Google Maps' \enquote{gold standard} travel times using those naïve times plus turns and traffic controls along the naïve routes. Validation reveals that this method yields accurate, free, and simple travel time predictions.

@@

    Travel time prediction underpins many transport planning processes and research problems. It is key to understanding urban accessibility, transport mode and route choice, and location decisions. Predicting congested travel times (i.e., when traffic congestion impedes flows) requires large volumes of real-time data on traffic conditions. Such data are proprietary, expensive, and beyond the reach of many planning practitioners and scholars. This is usually not a problem for well-resourced organizations with highly-skilled labor. However, technical expertise, complicated data and software requirements, and high costs pose insurmountable obstacles for many urban planning researchers, practitioners, and community advocates.

    Although typical \enquote{naïve} travel time prediction methods that minimize travel distance or street segment traversal time provide easy solutions, they ignore delays from turns and traffic controls and thus usually under-predicts real-world travel time. However, minimally-congested travel time prediction offers a middle ground of more-accurate predictions than naïve travel times do---yet much lower data and computing requirements than traffic-aware and congested travel time prediction. Focusing on driving, this study contributes a simple, accurate, and free method of minimally-congested travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this paper, across the Los Angeles urban area, our average trip's travel time prediction differs on average from Google's \enquote{gold standard} prediction by only 0.38 seconds. In contrast, simple---but common---naïve travel time prediction minimizing max-speed edge traversal times under-predicts Google's travel times by 182.71 seconds, on average.

    The rest of this paper is organized as follows. First it reviews the state of the art in travel time prediction and then considers the less-sophisticated methods that are standard in the current planning literature. Next it describes our data sources, models, and validation techniques. Then it presents our findings, highlighting both the feasibility and quality of our predictions in relation to simple naïve predictions or difficult expensive predictions. Finally, it concludes with a discussion of implications for research and practice.

@@

    Modern travel time prediction techniques tend to fall into one of two camps: 1) sophisticated, accurate, but difficult to execute; or 2) naïve, less-accurate, but simple to execute. The former is the state-of-the-art and has been the subject of much work in the recent computer science and engineering literatures. The latter tends to appear often in urban planning research and practice. Here we summarize the recent advances and current standards in these fields.

@@

    These advanced travel time prediction models require massive data input as they are usually trained on millions or billions of disaggregate empirical travel time records. For example, \citet{hou2018network} use 1.5 billion GPS-based travel time records in deep learning models to predict travel time in 13 road segments in St. Louis, yielding a MAPE of 6.8\%. Using more than 100,000 trips from Windows phone GPS data, \citet{woodard2017predicting}'s technique achieves a MAPE of 10.1\% when predicting travel times. \citet{pamula2023estimation} use three months of video sensing data at a 5-minute temporal resolution in a deep learning model to predict travel times in Poland, yielding a 6.8\% MAPE. Other studies use graph neural network models to predict travel times. For example, \citet{wang2023dynamic} use data from 100,000 Chinese ride-hailing trips in graph neural network and recurrent neural network models, yielding a MAPE of 15.4\%. \citet{vankdoth2023deep} achieve 3--4\% MAPEs using Q-Traffic and TaxiBJ data from Beijing and Chengdu in deep learning and graph neural network models.

@@

    Central to all of the preceding travel time prediction research is the concept of \textit{path solving}: identifying the shortest path between an origin and destination in a network model. In this case, the shortest path is defined as the path that minimizes travel time and could be realized methodologically in multiple ways. In the planning literature, studies variously minimize euclidean distance, network distance, network edge traversal time, free-flow travel time, or congested travel time. We discuss their trade-offs in this section.

    The actual path solving methods used in urban planning research and practice tend to differ substantially from the state of the art travel time prediction in the engineering and computer science literatures. The preceding section's models require extensive technical capacity, instrumentation, and data. In general, the planning literature's implementations of path solving tend to be simpler to execute, but potentially naïve and less-accurate. For example, many accessibility studies employ merely a (very) rough proxy for travel time by minimizing euclidean distance traveled instead \citep[e.g.,][]{macfarlane2021modeling,pearsall2020locating}. This offers the substantial benefit of simplicity (no network model, travel time date, or routing algorithms are needed) but offers a poor estimate of access in terms of travel time.

    Many other studies improve on this by instead solving shortest paths by minimizing network distance traveled. These studies often use data and tools from Esri, OpenStreetMap, or the US Census Bureau to measure network distances between origins and destinations \citep[e.g.,][]{mckenzie2020urban, jiao2021measuring, nicoletti2023disadvantaged, logan2019evaluating, tsou2005accessibility}. This offers the benefit of more-realistic distances between origins and destinations, but does not account for travel speeds. Other studies refine this by incorporating speed limit data into the network model to solve shortest paths by traversal time \citep{kuai2017examining,williams2020parks,he2020evaluating,salonen2013modelling,scott2008role,neutens2010equity,wang2013planning}. Traversal time is similar to---but distinct from---uncongested travel time. The former merely minimizes the sum of the ratio of street segment length and speed limit, whereas the latter accounts for stops, signals, and turns. Some researchers sidestep the challenges of needing a network model and travel speed data by using secondary traffic analysis zone (TAZ) travel time data aggregated by metropolitan planning agencies \citep[e.g.,][]{grengs2010intermetropolitan,yan2021toward,levine2012does}. These data offer simplicity and real-world empricism, but their coarse-grained zone-to-zone aggregation obfuscates accurate point-to-point travel time predictions.

    Finally, many planning scholars rely instead on the Google Maps API to obtain accurate origin-destination travel times and shortest paths \citep[e.g.,][]{fielbaum2021assessment,costa2021spatial, swayne2021integrating,hu2020estimating,cuervo2022dynamic,chen2020communities,hwang2024measuring}. As mentioned previously, Google travel times represent something of a \enquote{gold standard} in the literature, but come with drawbacks: the algorithms are closed-source and---for large batches of queries, such as for simulating metropolitan-scale trip taking---it can be expensive and require many API queries. Nevertheless, evidence shows Google travel times offer a good approximation of real-world travel times \citep{lin2021impact,fu2023comparative,alsobky2020estimating,wang2011estimating}.

@@

    This paper proposes a new generalizable method to solve this problem. In short, it collects a small free amount of data from the Google Maps API plus free open data to train a local model to predict accurate minimally-congested travel times using relatively simple standard tools and techniques. Figure~\ref{fig:workflow} shows a detailed workflow.

@@

    We define the study area as the convex hull around the intersection of the Los Angeles County boundary and the Los Angeles urban area boundary from the Global Human Settlement Layer's Urban Center Database \citep{florczyk2019description, GHS2019}. This allows us to retain the main urbanized area without adjacent metropolitan areas (such as the Inland Empire). We then model the drivable street network within this study area from OpenStreetMap using the OSMnx software \citep{boeing2017osmnx}, retaining the strongly connected component, to generate a graph with 782,825 nodes. This graph is unsimplified, meaning that nodes represent intersections and dead-ends, as well as curving street line geometry vertices. Table~\ref{tab:node_count} summarizes these nodes' tagged traffic control elements: these data are sparse and most intersections lack traffic control information.

@@

    Our final optimized model hyperparameterization uses 600 decision trees, random sampling with replacement in each decision tree, a maximum decision tree depth of 10, a non-limited maximum number of input explanatory features, a requirement of at least two samples to split at a decision node and one sample at a leaf node, and equal sample weighting. Finally, we check our trained model for overfitting by conducting 5-fold cross-validation across the whole sample using tuned hyperparameters: the resulting five MAE values (a, b, c, d, e) are all similar to each other, which indicates no overfitting problem as the hyperparameters are not overly specific to the training set.

@@

    Table~\ref{tab:validation_results} details our travel time prediction model's out-of-sample improvement over the initial naïve travel time calculation. Our model achieves a MAPE of 8.4\%---in line with the \textasciitilde10\% MAPEs seen in the state-of-the-art travel time prediction literature, but without their extensive and expensive input data requirements. In comparison, our initial naïve travel time calculation yields a much worse MAPE of 21.1\%. Whereas the naïve travel time calculation's MAE is 183.5 seconds from the Google travel time, our predicted travel times' MAE is just 75.2 seconds---an improvement by a factor of 2.4.

    \begin{table*}[tb!]
        \centering
        \caption{Results of out-of-sample (testing set) prediction of 1) our prediction model and 2) the initial naïve travel time calculation, validated against the corresponding Google travel times. The five validation measures are 1) MAPE, 2) MAE, 3) difference-in-means ($\delta$) and corresponding $t$-test $p$-value, 4) APR, and 5) $R^2$. See methods for definitions.}\label{tab:validation_results}
        \begin{tabular}{lrr}
            \toprule
            & Prediction model & Naïve calculation \\
            \midrule
            MAPE (\%)                     &             8.43 &             21.13 \\
            MAE (seconds)                 &            75.23 &            183.55 \\
            $\delta$ (seconds)            &             0.38 &           -182.71 \\
            $p$-value                     &             0.75 &             <0.01 \\
            APR                           &             1.01 &              0.79 \\
            $R^2$                         &             0.93 &              0.74 \\
            \midrule
            $n$                           &           8,272 &            8,272 \\
            \bottomrule
        \end{tabular}
    \end{table*}

@@

    \section{Discussion}
    Our study shows that our proposed method successfully meets its three intended criteria for middle-ground travel prediction: free, easy implementation, and high prediction accuracy. First, the model is built entirely on free and open-source data and software. It uses OpenStreetMap (OSM) for the street network and only a small amount of Google travel time data within the monthly free quota. In addition, its modeling process relies on open-source Python packages, including OSMnx for network routing and scikit-learn for training the random forest regression model. Second, this method is simple and easy to use. Rather than requiring sophisticated deep-learning techniques in the computer science literature, the proposed method only uses an out-of-the-box random forest regression model that can be implemented by anyone with basic experience in data science. Third, in terms of accuracy, our model's MAPE of 8.43\% substantially improves the MAPE of 21.13\% in the naïve model and falls within the MAPE range of 3\% to 17\% in the state-of-the-art models that we found in the recent literature. In summary, these results confirm that we proposed a free, easy-to-use, and accurate method for predicting minimally-congested travel time. It provides a middle-ground solution to improve travel time prediction accuracy without relying on expensive datasets and complicated methods.

    Importantly, these strengths are especially valuable for the audience that need them most---the urban planners. Although our proposed method does not aim to outperform the cutting-edge models, it provides an improved middle-ground solution for urban planners who lack access to expensive datasets and technical models typical in the computer science literature. Instead, they often rely on simple but less accurate methods, such as minimizing Euclidean distance, network distance, or edge traversal time based on the speed limit. These naïve methods, while easy to compute, systematically under-predict real-world travel times because they ignore the delays caused by turns and traffic controls. Our model addresses this problem by incorporating these features into the random forest regression model, which is much easier to implement than sophisticated deep learning models. In other words, our method does not replace the state-of-the-art methods for those who can implement them, but it makes meaningful progress over the naïve methods used by most urban planners.

    Building on these strengths, this study offers three contributions to planning research and practice. First, it provides a practical solution to bridge the methodological gap between overly complex and overly simple travel time predictions. On the one hand, accurate predictions of traffic-aware travel time require expensive, extensive, and often proprietary GPS data, which is often inaccessible to many urban planners. On the other hand, planners often use simple methods that under-predict real-world travel times. Our method provides a viable middle ground: a minimally congested travel time prediction method that is open, free, and easy-to-compute. Second, in addition to methodological accessibility, this model supports high-resolution, point-to-point travel time predictions across a large metropolitan area. Our case study is based on the populated area of Los Angeles County, demonstrating its applicability for metropolitan-scale travel time prediction. Third, the model is lightweight and efficient. It requires relatively short runtimes while maintaining prediction accuracy comparable to complex models. These contributions are important in urban planning practices because travel time can shift downstream planning processes based on accessibility, travel mode choice, and land use analysis. Inaccurate travel times can skew the results and cause misguided planning interventions. For example, based on under-predicted travel times, planners may falsely interpret high accessibility levels for marginalized neighborhoods far from employment centers. This could mask the real-world accessibility disadvantages and lead to underinvestment in these communities. Therefore, our proposed travel prediction method helps reduce transportation modeling bias and supports evidence-based planning interventions.

    Despite these contributions, our study has limitations that point future research directions. First, while used by billions of people around the world as the \enquote{best available} source of travel time prediction, Google travel time remains a black-box model based on users' GPS data. Future studies could use real-world observations, such as travel time from GPS traces, to train and validate the model. Second, our method is trained to predict minimally-congested travel time at 03:00 a.m. While it significantly improves upon naive travel time predictions, it does not reflect peak-hour traffic delays. Future studies could expand the model to different times of day to test its applicability. Additionally, our method predicts travel time but does not estimate the actual routes drivers may take. Given these limitations, it is not designed for cities with resources to develop sophisticated traffic models and extensive real-time urban informatics dashboards on travel demand and traffic flows. Rather, it serves urban planners in research and practice who lack access to such systems and need a low-cost technique for predicting better travel times than naïve methods. Third, our case study looked at the populated area in Los Angeles County, where the information in OpenStreetMap is relatively complete. To test the generalizability of this method, further studies should examine its performance in cities with sparser or lower-quality open data. Taken together, these limitations open opportunities to further improve travel time predictions across diverse geographical and temporal settings in the future.

@@

    Travel time prediction is central to questions of urban accessibility, mode and route choice, and individuals’ location decisions. True congested travel time vary drastically throughout the day and require large volumes of real-time or proprietary data, as well as complex algorithms, to model. These technical challenges and costs present a hurdle for many urban planners, who often turn to simpler models with lower and often free data and computing requirements. The traditional naïve methods are easy to implement, but produce inaccurate predictions.

    This study introduces a middle-ground method to simply but accurately predict minimally-congested travel time using free data. It improves on traditional, common, naïve predictions by incorporating a small one-off collection of free but proprietary high-quality Google data on travel times. Also of note is the sparseness of our open data: OpenStreetMap contains little information on traffic controls, as only around 28\% of the intersections in the study area have traffic control information. Even so, our prediction model using random forest regression still predicts Google travel times with high accuracy. This offers planning practitioners and scholars a better method to predict travel times for free with much better accuracy than traditional naïve methods offered.

    How can it be reused by others?
    Can it be extended to other times of day?
    Can it be extended to other study sites?

>>>>

\title{\papertitle}
\author[]{Redacted for review}%{\myname}
\affil[]{Redacted for review}%{\myaffiliation}
\date{}%{\paperdate}

@@

Travel time prediction is central to accessibility analyses, sustainable transportation infrastructure provision, and active transportation interventions. However, calculating accurate travel times, especially for driving, requires either extensive technical capacities and bespoke data, or sources like the Google Maps API that quickly becomes prohibitively expensive to analyze thousands or millions of trips necessary for metropolitan-scale planning. These obstacles make it difficult for researchers, practitioners, and community advocates to develop an evidence base for more sustainable transportation and equitable accessibility. This article develops a free, open-source minimally-congested driving time prediction model with minimal cost, data, and computational requirements. It trains and tests this model using the Los Angeles, California urban area as a case study by calculating naïve travel times then developing a predictive random forest regression model of Google Maps' \enquote{gold standard} travel times using those naïve times plus turns and traffic controls along the naïve routes. Validation reveals that this method yields accurate, free, and simple travel time predictions.

@@

Travel time prediction underpins many transport planning processes and research problems. It is key to understanding urban accessibility, transport mode and route choice, and location decisions. Predicting congested travel times (i.e., when traffic congestion impedes flows) requires large volumes of real-time data on traffic conditions. Such data are proprietary, expensive, and beyond the reach of many planning practitioners and scholars. This is usually not a problem for well-resourced organizations with highly-skilled labor. However, technical expertise, complex data and software requirements, and high costs pose insurmountable obstacles for many urban planning researchers, practitioners, and community advocates.

Although typical \enquote{naïve} travel time prediction methods (which minimize travel distance or street segment traversal time) provide easy solutions, they ignore delays from turns and traffic controls and thus usually under-predict real-world driving travel time. However, minimally-congested travel time prediction offers a middle ground of more-accurate predictions than naïve travel times do---yet much lower data and computing requirements than traffic-aware and congested travel time prediction.

To address this need for a middle ground, this article contributes a simple, accurate, and free method of minimally-congested driving travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this article, across the Los Angeles urban area, our average trip's travel time prediction differs on average from Google's \enquote{gold standard} prediction by only 0.38 seconds. In contrast, simple---but common---naïve travel time prediction minimizing max-speed edge traversal times under-predicts Google's travel times by 182.71 seconds, on average.

The rest of this article is organized as follows. First it reviews the state of the art in travel time prediction and then considers the less-sophisticated methods that are standard in the current planning literature. Next it describes our data sources, models, and validation techniques. Then it presents our findings, highlighting both the feasibility and quality of our predictions in relation to simple naïve predictions or difficult expensive predictions. Finally, it concludes with a discussion of implications for research and practice.

@@

Modern travel time prediction techniques tend to fall into one of two camps: (1) sophisticated, accurate, but difficult to execute; or (2) naïve, less-accurate, but simple to execute. The former is the state-of-the-art and has been the subject of much work in the recent computer science and engineering literatures. The latter tends to appear often in urban planning research and practice. Here we summarize the recent advances and current standards in these fields.

@@

These advanced travel time prediction models require massive data inputs as they are usually trained on millions or billions of disaggregate empirical travel time records. For example, \citet{hou2018network} use 1.5 billion GPS-based travel time records in deep learning models to predict travel time in 13 road segments in St. Louis, yielding a MAPE of 6.8\%. Using more than 100,000 trips from Windows phone GPS data, \citet{woodard2017predicting}'s technique achieves a MAPE of 10.1\% when predicting travel times. \citet{pamula2023estimation} use three months of video sensing data at a 5-minute temporal resolution in a deep learning model to predict travel times in Poland, yielding a 6.8\% MAPE.\@ Other studies use graph neural network models to predict travel times. For example, \citet{wang2023dynamic} use data from 100,000 Chinese ride-hailing trips in graph neural network and recurrent neural network models, yielding a MAPE of 15.4\%. \citet{vankdoth2023deep} achieve 3--4\% MAPEs using Q-Traffic and TaxiBJ data from Beijing and Chengdu in deep learning and graph neural network models.

@@

Central to all of the preceding travel time prediction research is the concept of \textit{path solving}: identifying the shortest path between an origin and destination in a spatial network model. In this case, the shortest path is defined as the path that minimizes travel time and could be realized methodologically in multiple ways. In the planning literature, studies variously minimize euclidean distance, network distance, network edge traversal time, free-flow travel time, or congested travel time. We discuss their trade-offs in this section.

The actual path solving methods used in urban planning research and practice tend to differ substantially from the state of the art travel time prediction in the engineering and computer science literatures. The preceding section's models require extensive technical capacity, instrumentation, and data. In general, the planning literature's implementations of path solving tend to be simpler to execute, but potentially naïve and less-accurate. For example, many accessibility studies employ merely a (very) rough proxy for travel time by minimizing euclidean distance traveled instead \citep[e.g.,][]{macfarlane2021modeling,pearsall2020locating}. This offers the substantial benefit of simplicity (no network model, travel time data, or routing algorithms are needed) but offers a poor estimate of access in terms of travel time.

Other studies improve on this by instead solving shortest paths by minimizing network distance traveled. These studies often use data and tools from Esri, OpenStreetMap, or the US Census Bureau to measure network distances between origins and destinations \citep[e.g.,][]{mckenzie2020urban, jiao2021measuring, nicoletti2023disadvantaged, logan2019evaluating, tsou2005accessibility}. This offers the benefit of more-realistic distances between origins and destinations, but does not account for travel speeds. Other studies refine this by incorporating speed limit data into the network model to solve shortest paths by traversal time \citep{kuai2017examining,williams2020parks,he2020evaluating,salonen2013modelling,scott2008role,neutens2010equity,wang2013planning}. Traversal time is similar to---but distinct from---uncongested travel time. The former merely minimizes the sum of the ratio of street segment length and speed limit, whereas the latter accounts for stops, signals, and turns. Some researchers sidestep the challenges of needing a network model and travel speed data by using secondary traffic analysis zone (TAZ) travel time data aggregated by metropolitan planning agencies \citep[e.g.,][]{grengs2010intermetropolitan,yan2021toward,levine2012does}. These data offer simplicity and real-world empricism, but their coarse-grained zone-to-zone aggregation obfuscates accurate point-to-point travel time predictions.

Finally, many planning scholars rely instead on the Google Maps API to obtain accurate origin-destination travel times and shortest paths \citep[e.g.,][]{fielbaum2021assessment,costa2021spatial, swayne2021integrating,hu2020estimating,cuervo2022dynamic,chen2020communities,hwang2024measuring}. As mentioned previously, Google travel times represent something of a \enquote{gold standard} in the literature, but come with drawbacks: the algorithms are closed-source and---for large batches of queries, such as for simulating metropolitan-scale trip taking---it can become prohibitively expensive and require many API queries. Nevertheless, evidence shows Google travel times offer a good approximation of real-world travel times \citep{lin2021impact,fu2023comparative,alsobky2020estimating,wang2011estimating}.

@@

This article proposes a new generalizable method to solve this problem. In short, it collects a small free amount of data from the Google Maps API plus free open data to train a local model to predict accurate minimally-congested travel times using relatively simple standard tools and techniques. Figure~\ref{fig:workflow} shows a detailed workflow.

@@

We define the study area as the convex hull around the intersection of the Los Angeles County boundary and the Los Angeles urban area boundary from the Global Human Settlement Layer's Urban Center Database \citep{florczyk2019description, GHS2019}. This allows us to retain the main urbanized area without adjacent metropolitan areas (such as the Inland Empire). We then model the drivable street network within this study area from OpenStreetMap using the OSMnx software \citep{boeing_modeling_2025}, retaining the strongly connected component, to generate a graph with 782,825 nodes. This graph is unsimplified, meaning that nodes represent intersections and dead-ends, as well as curving street line geometry vertices. Table~\ref{tab:node_count} summarizes these nodes' tagged traffic control elements: these data are sparse and most intersections lack traffic control information.

@@

Our final optimized model hyperparameterization uses 600 decision trees, random sampling with replacement in each decision tree, a maximum decision tree depth of 10, a non-limited maximum number of input explanatory features, a requirement of at least two samples to split at a decision node and one sample at a leaf node, and equal sample weighting. Finally, we check our trained model for overfitting by conducting 5-fold cross-validation across the whole sample using tuned hyperparameters: the resulting five MAE values (75.3, 74.5, 73.5, 73.3, 74.7) are all similar to each other, which indicates no overfitting problem as the hyperparameters are not overly specific to the training set.

@@

Table~\ref{tab:validation_results} details our travel time prediction model's out-of-sample improvement over the initial naïve travel time calculation. Our model achieves a MAPE of 8.4\%, in line with the approximately 10\% MAPEs seen in the state-of-the-art travel time prediction literature, but without their extensive and expensive input data requirements. In comparison, our initial naïve travel time calculation yields a much worse MAPE of 21.1\%. Whereas the naïve travel time calculation's MAE is 183.5 seconds from the Google travel time, our predicted travel times' MAE is just 75.2 seconds---an improvement by a factor of 2.4.

\begin{table*}[tb!]
    \centering
    \caption{Results of out-of-sample (testing set) prediction of (1) our prediction model and (2) the initial naïve travel time calculation, validated against the corresponding Google travel times. The five validation measures are (1) MAPE, (2) MAE, (3) difference-in-means ($\delta$) and corresponding $t$-test $p$-value, (4) APR, and (5) $R^2$. See methods for definitions.}\label{tab:validation_results}
    \begin{tabular}{lrr}
        \toprule
                                      & Prediction model & Naïve calculation \\
        \midrule
        MAPE (\%)                     &             8.43 &             21.13 \\
        MAE (seconds)                 &            75.23 &            183.55 \\
        $\delta$ (seconds)            &             0.38 &           -182.71 \\
        $p$-value                     &             0.75 &             <0.01 \\
        APR                           &             1.01 &              0.79 \\
        $R^2$                         &             0.93 &              0.74 \\
        \midrule
        $n$                           &           8,272 &            8,272 \\
        \bottomrule
    \end{tabular}
\end{table*}

@@

\section{Discussion}

Re-state the problem. This study set out to investigate if a middle-ground travel time prediction model was possible. It identified three criteria for success: (1) it should use free software and data and avoid bespoke sensor data collection; (2) it should be easy and accessible to use; (3) it should offer better accuracy than naïve predictions (e.g., minimizing Euclidean distance, network distance, street edge traversal time) and better spatial resolution than zone-to-zone travel times.

Our results show that our proposed method satisfies these criteria to offer an important middle-ground contribution. First, it relies only on open-source Python software, OpenStreetMap data, and a small free amount of Google travel time training data. Second, our model is easy to use: it does not require a complex computing environment suited for deep learning's extensive data and processing requirements. Rather, our model uses out-of-the-box tools that can be set up in seconds on a standard consumer-grade computer by anyone familiar with Python, the world's most popular programming language. Third, our model is accurate: its MAPE (8.4\%) is much lower than that of the naïve model and in line with state-of-the-art models (\textasciitilde3--17\%) with extensive data and computing requirements.

This middle ground is not necessarily for computer scientists, high-end labs with top-of-the-line computing hardware, or well-resourced agencies with extensive data budgets. Rather, it empowers the scholars and practitioners who need it the most---those working on the frontlines of urban planning and policymaking. These scholars and practitioners rely on simple but inaccurate routing models---such as minimizing Euclidean distance, network distance, or edge traversal time---when they lack the technical expertise, computing equipment, or extensive data needed to implement the literature's cutting-edge models. However, this study demonstrated how these naïve methods systematically under-predict real-world travel times. Our model addresses this problem by incorporating sparse open data on traffic control elements and turns into the prediction and achieves an accuracy in line with state-of-the-art methods---but without their data needs or computing requirements. In other words, our model does not replace the state-of-the-art methods for those with extensive resources available to implement them, but rather offers far more accurate predictions for urban planners who would otherwise fall back on simpler naïve models. Inaccurate travel times skew empirical understandings and mislead planning interventions. Our model represents an important contribution for urban planning scholarship and evidence-based practitioner interventions.

As demonstrated, this model supports high-resolution point-to-point travel time predictions across a large metropolitan area. Our implementation focused on Los Angeles but is not inherently tied to it, and future work should emphasize its generalizability in cities with potentially sparser or lower-quality open data. We trained this model as a proof-of-concept for minimally-congested times of day, but the same concept can be extended with similar training data for any other time of day. Future research should also expand the kinds of training and validation data. Although used by billions of people around the world as the public's \enquote{best available} source of travel time prediction, Google travel time remains a black box based on users' GPS data. Future work can use other empirical travel time data to further train and validate the model.

@@

Travel time prediction is central to questions of urban accessibility, mode and route choice, and individuals' location decisions. True congested travel time vary drastically throughout the day and require large volumes of real-time or proprietary data, as well as complex algorithms, to model. These technical challenges and costs present a hurdle for many urban planners, who often turn to simpler models with lower and often free data and computing requirements. The traditional naïve methods are easy to implement, but produce inaccurate predictions.

This article introduced a middle-ground method to simply but accurately predict minimally-congested travel time using free data. It improves on traditional, common, naïve predictions by training on a small one-off collection of free but proprietary high-quality Google data on travel times. Also of note is the sparseness of our open data: OpenStreetMap contains little information on traffic controls, as only around 28\% of the intersections in the study area have traffic control information. Even so, our prediction model using random forest regression still predicts Google travel times with high accuracy. This offers planning practitioners and scholars a better method to predict travel times for free with much better accuracy than traditional naïve methods offered.
